Tsbrowse: an interactive browser for Ancestral Recombination Graphs

### 2025-06-19 manuscript/revision/response_to_reviewers.tex

Summary: In this manuscript, the authors present tsbrowse, an open-source Python-based 
application for the interactive visualization of the components of Ancestral Recombination 
Graphs (ARGs). Because ARGs (in principle) contain all knowable information concerning 
coalescence and recombination in the history of a set of samples, there is great excitement 
in developing methods for inferring ARGs from empirical genomic sequence data, as well as 
for learning about evolutionary processes from patterns in inferred ARGs. A crucial component 
of both these endeavors is visualizing ARGs, which is difficult or unwieldy. Tsbrowse therefore 
represents an exciting contribution to the field which is likely to be of high utility and 
to be highly cited.

Opinion: This paper was straightforward, clear, and well written. I think it does a good job 
setting up the intellectual context for ARG-based research and advocating for the utility of 
tsbrowse. I had two minor, nitpicky comments, but overall, it was great. Thanks for a fun and 
easy review!\\
\textcolor{blue}{Thank you for your kind words!}

1. I think the order in which the supplemental figures are cited doesn’t match their numbering 
(i.e., S4 is cited before S1)\\
\textcolor{blue}{We have now rearranged the supplementary figures so that figure references 
appear in ascending order.}

2. In the discussion, the authors mention that methods for visualization of tree topology can 
typically only handle a few hundred nodes, but it might be worth doing a ``(but see…)'' for some 
of the approaches folks have taken for visualizing much larger trees (e.g., nextstrain, onezoom).\\
\textcolor{blue}{We had already included a citation to the OneZoom project (and the Treenome browser) in the 
sentence that followed the one in question "… Visualisation of large-scale tree topologies with millions of nodes 
requires much more sophisticated approaches to capture topological features at different scales, 
and is an active research area". Thanks for bringing Nextstrain to our 
attention, and we have now included \href{{Wong et al 2024}}{Hadfield et al.} as a citation for visualisation of 
medium-sized toplogies.}\\

@@

\textbf{Reviewer: 2}\\

Comments to the Author\\

I have only a few minor comments:\\
1. ``and storing as a compressed .tsbrowse file. (The .tsbrowse file is also a valid input for
tszip, a general utility for compressing tskit files.)'' Sorry, but I don't understand this sentence. 
More precisely, both the compressed and the uncompressed files have the .tbrowse extension? If yes, 
that's a weird choice and I would like to see some justification. You should also make clear if tszip 
is a part of tskit or of twbrowse.\\
\textcolor{blue}{To clarify, tszip is part of the tskit ecosystem. It is not part of the tsbrowse 
package but can be used to compress the tables output from the tsbrowse preprocess step. Thank you 
for pointing out this potentially unclear statement. We have now deleted the text 
in brackets, which we believe isn't crucial for understanding the data model.}\\

@@

2. ``web-browser'' - web browser\\
\textcolor{blue}{Fixed.}\\

3. ``Fig S4, for example, shows a screenshot of tsbrowse''. This is a matter of personal taste, but I do
not like to refer to a Figure in the supplementary materia just as if it were in the article; if the 
Figure is relevant, it should be in the article. If it's not relevant, that you should refer to it without
hinting that it is important.\\
\textcolor{blue}{We agree with your opinion, and believe that this figure is important 
to demonstrate the application of tsbrowse to large-scale datasets (which is a key design 
feature of the app). However, given the journal guidelines for page limit, we are unable 
to move this figure to the main matter.}\\

4. Looking at the github repo, I can find the installation instruction using pip only. In my experience 
pip is quite picky. I strongly suggest you to provide tsbrowse via conda-forge/bio-conda or a containerized 
solution (e.g. docker).\\
\textcolor{blue}{We have now created a Docker container in order to ensure a smooth user experience across 
computing environments. Usage instructions have been added to the app documentation.}\\

@@

\textbf{Reviewer: 3}\\

Comments to the Author\\
This paper offers a browser-based visualisation tool for ancestral recombination graphs (ARGs), which is 
a data structure representing gene genealogies for a sample of DNA sequences. The complication in the 
presence of recombination is that these structures are not simple trees, but - because of the possibility 
of recombination - graphs with reticulated nodes, which makes exploratory data analysis difficult.

@@

I agree with the authors that there is currently a lack of visualisation tools for ARGs, and I expect the 
current paper to provide a useful addition to the catalog of tools for ARG simulation and inference. Tsbrowse 
integrates with the widely used tskit ecosystem of software tools, developed by some of the current authors, 
which will facilitate a rapid uptake. The authors make a compelling argument for the need for visualisation 
tools - Fig S1 in particular shows that four different and popular tools for ARG inference are giving 
qualitatively completely different insights into their reconstructed ARGs (incidentally, it's a shame that 
that figure couldn't make it into the main paper, though I recognise that this journal has strict page limits).\\
\textcolor{blue}{Thank you for your valuable feedback on the manuscript, and on the utility of tsbrowse!}

\textcolor{blue}{We completely agree that this figure would have been a compelling main display item had the journal 
page/figure limits allowed for it. Considering the importance of the existing main display items in illustrating 
the core aspects of tsbrowse, we have decided to retain this figure in the supplementary.}\\

My main comment is that tsbrowse does not actually seem to be an ARG visualisation tool. It is a useful way 
to visualise mutations, tree-sequence edges, and summary statistics, but nowhere will you actually see a tree 
(much less an ARG). On rereading the paper I see that the manuscript has been carefully worded not to include 
such a claim, but a casual reader might expect tsbrowse to be able to do so, as other tools such as IcyTree can. 
I do get fed up with reviewers tossing additional requests on top of an already polished piece of work, but 
in this case I think it would be a helpful addition! - would it really be too much work for tsbrowse to be 
able to draw at least the local trees? (For researchers with smaller datasets, being able to interrogate the 
fine details of an ARG would be equally valuable as the complementary selling-points around scalability.) At 
the very least I would like to see the manuscript at least be a little more up-front about this omission.\\
\textcolor{blue}{This is an important observation, and we agree that drawing (parts of) local trees would be 
a great feature. For smaller datasets, tools such as 
\href{https://pypi.org/project/tskit-arg-visualizer/}{tskit-arg-visualizer} and existing tskit infrastructure, 
for example functions \texttt{draw\_text()} and \texttt{draw\_svg()} already provide this functionality. A key 
motivation for this work is to guide large-scale inference, and implementing tree visualisation that is 
interpretable at this scale would require significant thought towards design, as well as considerable effort. 
We therefore think that implementing this feature is not feasible for this work and is better scoped as future 
work or as a separate endeavour. In line with this, we have already explicitly stated in the Discussion that drawing 
local trees is an important future direction. In the text, we do state that tsbrowse "provides visualisations of 
the information structure of ARGs", which we hope implies that it is a visualiser for the \textit{properties} 
of the ARG, rather than its topology itself.}\\

@@

Minor comments:\\

Manuscript p1: 'it is only with recent breakthroughs in inference methods...that practical application 
has become possible'. I appreciate the need to emphasise the recent increase in interest in inference for ARGs, 
but I think this is selling older methods a little short. There are ARG inference methods going back at least 
as far as Griffiths and Marjoram (1996) 'Ancestral Inference from Samples of DNA Sequences' JCB 3(4):479-502. 
True they are not scalable in the manner of recent methods, but they *are* practical applications, and they have 
the advantage that they actually sample from the claimed model without introducing any model heuristics.\\
\textcolor{blue}{Thank you for pointing this out, and you are right that we are somewhat overstating the 
case for recent methods. We have changed the word "practical" to "widespread", which we feel is an uncontroversial 
statement. The history of ARG inference is a complicated and confusing subject, dealt with exhaustively by 
\href{https://academic.oup.com/genetics/article/228/1/iyae100/7714980}{Wong et al. 2024}, and would be impossible to accurately summarise in the available space. We note, though, 
that two of the recent methods cited here (ARGweaver and SINGER) sample ARGs from the posterior defined by a 
coalescent model in a statistically rigorous manner.}\\

@@

Manuscript p1: Typo 'developements'.\\
\textcolor{blue}{This is now fixed.}\\

Manuscript p2: Where you mention 'Bokeh', can you clarify what this is (e.g. with a citation)?\\
\textcolor{blue}{Yes, Bokeh is a Python library for interactive visualisation and a citation to it 
has been added.}\\

Supplement p1, Sec 1.4: ``The ARG shown in Fig 4'' should say "The ARG shown in Suppl. Fig 4" ?\\
\textcolor{blue}{This is now fixed.}\\

Fig S2 caption - 'Consortium (2015)' is citing incorrectly.\\
\textcolor{blue}{Fixed (removed this citation as it was not required).}\\

@@

Installation notes:

I had a few problems getting tsbrowse to work. There are two different sets of installation instructions:

1. https://pypi.org/project/tsbrowse/
2. https://tskit.dev/tsbrowse/docs/stable/contributing.html

I use a Mac+Homebrew+Chrome and Option 2 seems more fitting. ("python -m pip install tsbrowse" as in 
option 1 will fail because Homebrew's Python environment is externally managed.)\\
\textcolor{blue}{We apologise for the confusion surrounding installation notes. The recommended way to install 
tsbrowse is via pip within a virtual environment. We have fixed the documentation so that instructions 
to install tsbrowse as a new user are clearly stated at the beginning of the documentation,
and are separate from developer instructions.}\\

@@

Bug(?):\\

I frequently had problems rendering the window in the 'Mutations' tab, which tended to get stuck at the 
circling wheel. When clicking the 'Mutations' button, my terminal returned the following error:

2025-05-27 12:29:41,646 [5441] ERROR    tornado.application: Exception in callback functools.partial(>, .wrapped() 
done, defined at \\
/Users/reviewer/myenv/lib/python3.13/site-packages/panel/io/server.py:153> 
exception=AbbreviatedException(, ValueError('Supplied cmap winter not found among matplotlib, bokeh, or colorcet 
colormaps.'), )>)\\
\textcolor{blue}{Thank you for bringing this bug to our attention. We have solved the issue and released an 
updated version of the package.}\\

### 2025-06-19 manuscript/paper.tex
(hopefully) final readthrough before submitting revision

\abstract{
    %\textbf{Motivation:}
    Ancestral Recombination Graphs (ARGs) represent the interwoven paths of genetic
    ancestry for a set of recombining sequences. The ability to capture the
    evolutionary history of samples makes ARGs valuable in a wide range of
    applications in population and statistical genetics. ARG-based approaches are
    increasingly becoming a part of genetic data analysis pipelines due to
    breakthroughs enabling ARG inference at biobank-scale. However, there is a lack
    of visualisation tools, which are crucial for validating inferences and
    generating hypotheses.
    %\textbf{Results:} 
    We present \texttt{tsbrowse}, an open-source Python web-app for the interactive
    visualisation of the fundamental building-blocks of ARGs, i.e.,  nodes, edges
    and mutations. We demonstrate the application of \texttt{tsbrowse} to various
    data sources and scenarios, and highlight its key features of browsability
    along the genome, user interactivity, and scalability to very large sample
    sizes.\\ \textbf{Availability:}\\ Python package:
    \texttt{\url{https://pypi.org/project/tsbrowse/}}, \\ Development version:
    \texttt{\url{https://github.com/tskit-dev/tsbrowse}}, \\ Documentation:
    \texttt{\url{https://tskit.dev/tsbrowse/docs/}}, \\ DOI: 
    \texttt{\url{https://doi.org/10.5281/zenodo.15683039}}\\
    %\textbf{Contact:} \href{name@email.com}{name@email.com}\\
    %\textbf{Supplementary information:} Available at
    %\textit{\textcolor{red}{\href{https://www.google.com}{Bioinformatics}
    %online}.}}
}

@@

\subsection{Architecture}
\texttt{Tsbrowse} is a modular client-server application
written in Python, optimised for ARGs with millions
of samples (Fig.~\ref{fig:Figure_1}).
The application has two basic commands: \texttt{preprocess}
and \texttt{serve}. The \texttt{preprocess} command
takes an input ARG file and augments the \texttt{tskit}
tables with additional columns,
precomputing all the information required for visualisation,
and storing as a compressed \texttt{.tsbrowse} file.
%(The \texttt{.tsbrowse} file is also a valid input for
%\texttt{tszip}, a general utility for compressing \texttt{tskit}
%files.)
To visualise an ARG we then run \texttt{tsbrowse serve}, which
by default will open a browser window on the local machine, but
also supports running the server on a remote machine across the
network.
This client-server architecture has some important advantages over
a monolithic single-machine approach. Most importantly,
the ARG being visualised remains in-situ and does not need to be
downloaded from the server.

@@

\subsection{User Interface} \label{subsec:User_Interface} \begin{figure}
    \centering \includegraphics[width=0.95\linewidth]{figures/MainFig2.png}
    \caption{Screenshot of the Edges tab in \texttt{tsbrowse}.
        Visualisation of the edges in a simulated ARG with a strong selective
        sweep in the middle of the genome (see Supplement for details).
        Edges are shown in the main browser pane on the left, with additional
        histograms summarising the edges on the right.
        The effect of the sweep can be seen by the lack of edges crossing the
        centre of the simulated chromosome due to an excess of recent coalescent
        events. Moving away from the focal site, the oldest edges are the
        first to rejoin the ARG, followed by more recent edges, resulting in a
        wedge-like pattern of missing edges in the centre.
        \label{fig:Figure_2}}
\end{figure}
% \texttt{Tsbrowse}'s interface is built on the \texttt{Holoviz} ecosystem
% \citep{Holoviz}, which brings together \texttt{Datashader}, \texttt{Bokeh},
% \texttt{Panel}, \texttt{Holoviews} and other libraries to create seamless data
% visualisation pipelines even for large datasets. A key challenge with
% large-scale data visualisation is to maintain a responsive user interface.
% \texttt{Datashader} aggregates data and renders only the necessary data points
% at a given zoom level, thereby minimising the computational load on the
% browser. A \texttt{Bokeh} plotting back-end provides useful tools and widgets.
% The \textit{pan}, \textit{select} and \textit{zoom} tools enable interactivity
% with individual data points, allowing users to gain deeper insights from the
% data. The \texttt{serve} step accepts an optional input file containing
% sequence annotations. This allows users to overlay contextual information about
% genes or other sequence features. 
The user interface is presented as a dashboard, with pages to
describe various aspects of the ARG.
Fig.~\ref{fig:Figure_2}
demonstrates the user interface using the Edges page as an example.
The plot on the left is an overview of the 33,929 edges
in a simulated ARG with a strong selective sweep (see
Supplement for details).
Each edge is depicted as a horizontal line connecting the genomic coordinates
of the
parent and child nodes on the x-axis, and the y-axis shows the time of
either the parent or the child, as chosen by the user.
The user can interact with the main ``genome browser'' window
using a set of controls on the top-right corner provided by 
Bokeh~\citep{bokeh},allowing them to pan and zoom as required. 
The histograms to the
right then summarise the edges as depicted in the browser window.
A similar browser interface is provided for
mutations
%(e.g. Fig~\ref{fig:Supplementary_Figure_1})
(e.g., Fig.~S1)
and nodes
%(e.g. Fig~\ref{fig:Supplementary_Figure_2}).
(e.g., Fig.~S2).
\texttt{Tsbrowse} also provides an interactive table viewer with flexible
searching and sorting utilities, which is a valuable debugging utility for
developers.

@@

\subsection{Applications}
The purpose of \texttt{tsbrowse} is to provide an interactive view of the
\texttt{tskit} ARG data model to guide intuition, improve inference quality
control and facilitate debugging. An example of how we can deepen our understanding
using the genome-browser-like perspective provided by \texttt{tsbrowse}
is given in the simulated ARG of Fig.~\ref{fig:Figure_2}, where
the gap in the edges view corresponds to the characteristic dip in diversity
of a selective sweep. Comparing this ground truth to the ARGs inferred
%by four different inference methods in Fig~\ref{fig:Supplementary_Figure_3},
by four different inference methods in Fig.~S3,
we can see that there are substantial qualitative differences between the
results. These differences in ancestral haplotypes illustrated by
the edges view are unlikely to be captured by the tree-by-tree distance
metrics usually used to evaluate
inference methods~\citep[e.g.][]{kelleher2019inferring,zhang2023biobank},
providing further motivation for new and improved ways to compare
simulated and inferred ARGs~\citep{fritze2024forest}.

Interest in ARG-based methods is burgeoning, but the methods are new
and practical guidance on applying inferences to real data is lacking.
Data filtering is essential, and the effects of the
choices that must be made along any bioinformatics pipeline on the final
ARG are hard to predict and quantify. \texttt{Tsbrowse} was primarily
developed as a way to quickly visualise the effects of such filtering
choices on ARGs inferred by \texttt{tsinfer}, and it is now an indispensable
%element of the inference pipeline. Fig~\ref{fig:Supplementary_Figure_4} 
element of the inference pipeline. Fig.~S4
shows a region of the 1000G data with gaps in site density that are spanned by
 exceptionally long edges, which are likely to bias downstream statistics.
These interactive visualisations have also helped diagnose issues with
the \texttt{tsinfer} inference algorithm.
%Fig~\ref{fig:Supplementary_Figure_2} shows the genomic spans of
Fig.~S2 shows the genomic spans of
ancestral nodes in an ARG inferred with \texttt{tsinfer}, demonstrating a clear
excess of long haplotypes in the very ancient past. These insights
have helped guide development and
may lead to significant improvements in performance.

@@

\section{Discussion} \label{sec:Discussion}
Visualisation of tree topology is a central task in phylogenetics,
and although numerous tools
exist~\citep[e.g.][]{huson2007dendroscope,vaughan2017icytree},
the methods can typically only handle a few hundred nodes 
(but see e.g., \citet{hadfield2018nextstrain})
Visualisation of large-scale tree topologies with millions of nodes
requires much more sophisticated approaches
to capture topological features at different scales,
and is an active research area~\citep{wong2022dynamic,kramer2023treenome}.
Adapting such methods, and integrating into \texttt{tsbrowse} to provide
a local tree viewer that operates at the million-node scale
is an important direction for future work.
An interactive viewer for the entire ARG topology, capturing
semantic properties of the graph at a range of scales, is an even more
ambitious goal, and would be a major asset for the field.

>>>>

\abstract{
    %\textbf{Motivation:}
    Ancestral Recombination Graphs (ARGs) represent the interwoven paths of genetic
    ancestry of a set of recombining sequences. The ability to capture the
    evolutionary history of samples makes ARGs valuable in a wide range of
    applications in population and statistical genetics. ARG-based approaches are
    increasingly becoming a part of genetic data analysis pipelines due to
    breakthroughs enabling ARG inference at biobank-scale. However, there is a lack
    of visualisation tools, which are crucial for validating inferences and
    generating hypotheses.
    %\textbf{Results:} 
    We present \texttt{tsbrowse}, an open-source Python web-app for the interactive
    visualisation of the fundamental building-blocks of ARGs, i.e.,  nodes, edges
    and mutations. We demonstrate the application of \texttt{tsbrowse} to various
    data sources and scenarios, and highlight its key features of browsability
    along the genome, user interactivity, and scalability to very large sample
    sizes.\\ \textbf{Availability:}\\ Python package:
    \texttt{\url{https://pypi.org/project/tsbrowse/}}, \\ Development version:
    \texttt{\url{https://github.com/tskit-dev/tsbrowse}}, \\ Documentation:
    \texttt{\url{https://tskit.dev/tsbrowse/docs/}}, \\ DOI: 
    \texttt{\url{https://doi.org/10.5281/zenodo.15683039}}\\
    %\textbf{Contact:} \href{name@email.com}{name@email.com}\\
    %\textbf{Supplementary information:} Available at
    %\textit{\textcolor{red}{\href{https://www.google.com}{Bioinformatics}
    %online}.}}
}

@@

\subsection{Architecture}
\texttt{Tsbrowse} is a modular client-server application
written in Python, optimised for ARGs with millions
of samples (Fig.~\ref{fig:Figure_1}).
The application has two basic commands: \texttt{preprocess}
and \texttt{serve}. The \texttt{preprocess} command
takes an input ARG file and augments the \texttt{tskit}
tables with additional columns,
precomputing all the information required for visualisation,
and storing it as a compressed \texttt{.tsbrowse} file.
%(The \texttt{.tsbrowse} file is also a valid input for
%\texttt{tszip}, a general utility for compressing \texttt{tskit}
%files.)
To visualise an ARG we then run \texttt{tsbrowse serve}, which
by default will open a browser window on the local machine, but
also supports running the server on a remote machine across the
network.
This client-server architecture has some important advantages over
a monolithic single-machine approach. Most importantly,
the ARG being visualised remains in-situ and does not need to be
downloaded from the server.

@@

\subsection{User Interface} \label{subsec:User_Interface} \begin{figure}
    \centering \includegraphics[width=0.95\linewidth]{figures/MainFig2.png}
    \caption{Screenshot of the Edges view in \texttt{tsbrowse}.
        Visualisation of the edges in a simulated ARG with a strong selective
        sweep in the middle of the genome (see Supplement for details).
        Edges are shown in the main browser pane on the left, with additional
        histograms summarising the edges on the right.
        The effect of the sweep can be seen by the lack of edges crossing the
        centre of the simulated chromosome due to an excess of recent coalescent
        events. Moving away from the focal site, the oldest edges are the
        first to rejoin the ARG, followed by more recent edges, resulting in a
        wedge-like pattern of missing edges in the centre.
        \label{fig:Figure_2}}
\end{figure}
% \texttt{Tsbrowse}'s interface is built on the \texttt{Holoviz} ecosystem
% \citep{Holoviz}, which brings together \texttt{Datashader}, \texttt{Bokeh},
% \texttt{Panel}, \texttt{Holoviews} and other libraries to create seamless data
% visualisation pipelines even for large datasets. A key challenge with
% large-scale data visualisation is to maintain a responsive user interface.
% \texttt{Datashader} aggregates data and renders only the necessary data points
% at a given zoom level, thereby minimising the computational load on the
% browser. A \texttt{Bokeh} plotting back-end provides useful tools and widgets.
% The \textit{pan}, \textit{select} and \textit{zoom} tools enable interactivity
% with individual data points, allowing users to gain deeper insights from the
% data. The \texttt{serve} step accepts an optional input file containing
% sequence annotations. This allows users to overlay contextual information about
% genes or other sequence features. 
The user interface is presented as a dashboard, with views to
describe various aspects of the ARG.
Fig.~\ref{fig:Figure_2}
demonstrates the user interface using the Edges view as an example.
The plot on the left is an overview of the 33,929 edges
in a simulated ARG with a strong selective sweep (see
Supplement for details).
Each edge in the ARG is depicted as a horizontal line connecting the genomic coordinates
of the
parent and child nodes on the x-axis, and the y-axis shows the time of
either the parent or the child, as chosen by the user.
The user can interact with the main ``genome browser'' window
using a set of controls on the top-right corner provided by 
Bokeh~\citep{bokeh}, allowing them to pan and zoom as required. 
The histograms to the
right then summarise the edges as depicted in the browser window.
A similar browser interface is provided for
mutations
%(e.g. Fig~\ref{fig:Supplementary_Figure_1})
(e.g., Fig.~S1)
and nodes
%(e.g. Fig~\ref{fig:Supplementary_Figure_2}).
(e.g., Fig.~S2).
\texttt{Tsbrowse} also provides an interactive table viewer with flexible
searching and sorting utilities, which is a valuable debugging utility for
developers.

@@

\subsection{Applications}
The purpose of \texttt{tsbrowse} is to provide an interactive view of the
\texttt{tskit} ARG data model to guide intuition, improve inference quality
control and facilitate debugging. An example of how we can deepen our understanding
using the genome-browser-like perspective provided by \texttt{tsbrowse}
is given in the simulated ARG of Fig.~\ref{fig:Figure_2}, where
the gap in the Edges view corresponds to the characteristic dip in diversity
of a selective sweep. Comparing this ground truth to the ARGs inferred
%by four different inference methods in Fig~\ref{fig:Supplementary_Figure_3},
by four different inference methods in Fig.~S3,
we can see that there are substantial qualitative differences between the
results. These differences in ancestral haplotypes illustrated by
the Edges view are unlikely to be captured by the tree-by-tree distance
metrics usually used to evaluate
inference methods~\citep[e.g.][]{kelleher2019inferring,zhang2023biobank},
providing further motivation for new and improved ways to compare
simulated and inferred ARGs~\citep{fritze2024forest}.

Interest in ARG-based methods is burgeoning, but the methods are new
and practical guidance on applying inferences to real data is lacking.
Data filtering is essential, and the effects of the
choices that must be made along any bioinformatics pipeline on the final
ARG are hard to predict and quantify. \texttt{Tsbrowse} was primarily
developed as a way to quickly visualise the effects of such filtering
choices on ARGs inferred by \texttt{tsinfer}, and it is now an indispensable
%element of the inference pipeline. Fig~\ref{fig:Supplementary_Figure_4} 
element of the inference pipeline. Fig.~S4
shows a region of the 1000 Genomes data with gaps in site density that are spanned by
 exceptionally long edges, which are likely to bias downstream statistics.
These interactive visualisations have also helped diagnose issues with
the \texttt{tsinfer} inference algorithm.
%Fig~\ref{fig:Supplementary_Figure_2} shows the genomic spans of
Fig.~S2 shows the genomic spans of
ancestral nodes in an ARG inferred with \texttt{tsinfer}, demonstrating a clear
excess of long haplotypes in the very ancient past. These insights
have helped guide development and
may lead to significant improvements in performance.

@@

\section{Discussion} \label{sec:Discussion}
Visualisation of tree topology is a central task in phylogenetics,
and although numerous tools
exist~\citep[e.g.][]{huson2007dendroscope,vaughan2017icytree},
the methods can typically only handle a few hundred nodes 
(but see e.g., \citet{hadfield2018nextstrain}).
Visualisation of large-scale tree topologies with millions of nodes
requires much more sophisticated approaches
to capture topological features at different scales,
and is an active research area~\citep{wong2022dynamic,kramer2023treenome}.
Adapting such methods, and integrating them into \texttt{tsbrowse} to provide
a local tree viewer that operates at the million-node scale
is an important direction for future work.
An interactive viewer for the entire ARG topology, capturing
semantic properties of the graph at a range of scales, is an even more
ambitious goal, and would be a major asset for the field.
